---
title: "AXIELL R HACKS"
subtitle: "Uzitecne prikazy a funkce pro migrace dat"
author: "Kateřina Křížová"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

```{r include = F}
knitr::opts_chunk$set(echo = F, eval = F)
```

\newpage 

# *GENERAL*

## I M P O R T / E X P O R T

### TXT

read.table = [function](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/read.table)
read.csv = [function](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/read.table)
read_csv = [function]()

~ ENCODING ~

fileEncoding="cp1250"
encoding="utf-8"

#### Read txt

```{r}
df <- read.table("D:/VURV/R_hacks/data/simple_table_4read.txt", header = TRUE, sep = ";")
```

#### Read csv

read.csv : sep is always "," 

```{r}
df <- read.csv("D:/VURV/R_hacks/data/simple_table_4read.csv", header = TRUE, sep = ";")
```

#### Write to Clipboard

```{r}
write.table(df, "clipboard", sep = "\t", row.names = FALSE)
```


#### Memory issues

\#Error in .jcall(cell, "V", "setCellValue", value) :  java.lang.OutOfMemoryError: GC overhead limit exceeded

```{r CHATGPT}

# Determine the number of rows and split point
total_rows <- nrow(APND_pk_rus)
split_points <- c(round(total_rows / 3), 2 * round(total_rows / 3))

# Split the dataframe into two parts
part1 <- APND_pk_rus[1:split_points[1], ]
part2 <- APND_pk_rus[(split_points[1] + 1):split_points[2], ]
part3 <- APND_pk_rus[(split_points[2] + 1):total_rows, ]

# Save each part to Excel files
gc()
write.xlsx(part1, file = paste0(path_ces, proj, "/", proj, "_PK_R_part1.xlsx"))
gc()
write.xlsx(part2, file = paste0(path_ces, proj, "/", proj, "_PK_R_part2.xlsx"))
gc()
write.xlsx(part3, file = paste0(path_ces, proj, "/", proj, "_PK_R_part3.xlsx"))

```



### XLSX

 ~ PACKAGES ~
 
 readxl, xlsx, openxlsx
 
 
 ~ FUNCTIONS ~
 
 readxl::read_excel()
 xlsx::read.xlsx()
 
 xlsx::write.xlsx()

#### Read excel

```{r}

df <- read_excel("D:/VURV/R_hacks/data/simple_table_4read.xlsx" , sheet = 2, header = TRUE)
```


ISSUE: 'Warning: Expecting numeric in E1251 / R1251C5: got '41 (cm)'  

SOLUTION: df <- read_excel(paste0(path_snd, "muzbe_rozmery_ptaci.xlsx"), sheet = 1, **col_types="text"**)

#### Read multiple excel sheets

potrebny balik "data.table"

```{r}
read_excel_allsheets <- function(filename, tibble = FALSE) {
    # I prefer straight data.frames
    # but if you like tidyverse tibbles (the default with read_excel)
    # then just pass tibble = TRUE
    sheets <- readxl::excel_sheets(filename)
    x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X, col_types = "text"))
    if(!tibble) x <- lapply(x, as.data.frame)
    names(x) <- sheets
    x
}

mysheets <- read_excel_allsheets(paste0(path_proj, "20xx.xlsx")) # specifikuj excel file
# list_data <- Map(as.data.frame, mysheets) # nesjpis neni nutny krok
df <- rbindlist(mysheets) # pripadne list_data
```


#### Write excel

```{r}
write.xlsx(x = ToothGrowth,                       
           file = "ToothGrowth.xlsx")
```

#### Write excel / more dataframes to sheets

```{r}
exi <- paste0(path_ces, proj, "/", proj, "_KAT.xlsx")

if(file.exists(exi)){
  print("File already exists !!!")
} else {
  f2s = exi
  write.xlsx(as.data.frame(SAVE1), file= f2s, sheetName="sparovano", row.names=FALSE)
  write.xlsx(as.data.frame(SAVE2), file= f2s, sheetName="duplicity", append=TRUE, row.names=FALSE)
  write.xlsx(as.data.frame(SAVE3), file= f2s, sheetName="nesparovanoMUS", append=TRUE, row.names=FALSE)
  write.xlsx(as.data.frame(SAVE4), file= f2s, sheetName="nesparovanoCES", append=TRUE, row.names=FALSE)
  write.xlsx(as.data.frame(SAVE0), file= f2s, sheetName="puv. vypis CES", append=TRUE, row.names=FALSE)
  rm(exi)
  rm(f2s)
  }
```

#### Write excel / levels to sheets

```{r}

install.packages("writexl")
library(writexl)

empty_data <- data.frame()
write_xlsx(empty_data, path = paste0(path_proj, "vystupy/", proj, "_PK_NV.xlsx")) 
  
for (i in Lzdroj) {
  subset_df <- save %>% filter(REL_zdroj == i)
  shtnam <- i
  write.xlsx(as.data.frame(subset_df), file = paste0(path_proj, "vystupy/", proj, "_PK_NV.xlsx"), 
             sheetName = i, row.names = FALSE, append = TRUE)
}
 
```


#### Write excel - memory issue

Error  
pouzit funkci gc() = Grabage Collection -> vycisti pamet po ulozeni predchoziho listu  

```{r}

exi <- paste0(path_ces, proj, "/", proj, "_KAT_PVO.xlsx")

if(file.exists(exi)){
  print("File already exists !!!")
} else {
  f2s = exi
  write.xlsx(as.data.frame(SAVE1), file= f2s, sheetName="sparovano", row.names=F, showNA = F)
  gc()
  write.xlsx(as.data.frame(SAVE2), file= f2s, sheetName="duplicity", append=T, row.names=F, showNA = F)
  gc()
  write.xlsx(as.data.frame(SAVE3), file= f2s, sheetName="nesparovanoMUS", append=T, row.names=F, showNA = F)
  gc()
  write.xlsx(as.data.frame(SAVE4), file= f2s, sheetName="nesparovanoCES", append=T, row.names=F, showNA = F)
  gc()
  write.xlsx(as.data.frame(SAVE0), file= f2s, sheetName="puv. vypis CES", append=T, row.names=F, showNA = F)
  rm(exi)
  rm(f2s)
  }
```

### CSV SAVE

#### Conditional SAVE


```{r}

f2s <- paste0(path_kk, proj, "_allXLSX.csv")

if(file.exists(f2s)){
  print("You've been here before.")
  rm(f2s)
} else {
  write.table(df_filled, file = f2s, 
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="cp1250")
  rm(f2s)
  print("File saved!")
  }
```

#### Factor save to multiple CSV

```{r}

mus_tab$rok12 <- as.factor(mus_tab$rok12)

df_list <- split(mus_tab, mus_tab$rok12)

for (i in seq_along(df_list)) {
  subset_df <- df_list[[i]]
  group_name <- unique(subset_df$rok12)
  # write.csv(subset_df, paste0(path_kk, "import/", proj, "_import_", group_name, ".csv"), row.names = FALSE)
  write.table(subset_df, file = paste0(path_kk, "import/roky/", proj, "_import_", group_name, ".csv"),
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="UTF-8")
}

```

#### Sample (random)

```{r}

smpl <- paste0(imp, "_sample")

sam <- sample_n(mus_tab, 10)

f2s <- paste0(path_kk, "import/", proj, "_import_", smpl, ".csv")

if(file.exists(f2s)){
  print("You've been here before!")
  rm(f2s)
} else {
  write.table(sam, file = f2s,
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="UTF-8")
            # na = "", fileEncoding="cp1250")
  print("Import csv  W R I T T E N  !")
  rm(f2s)
}
```


## N U M E R I C S

### Leading zeros -add

```{r LEADING ZEROS}

# package
require(stringr) # ::str_pad

# command
mutate(MPORC = str_pad(PorC_S, 6, pad = "0"))
```

mutate = create a new column  
MPORC = name of a new column
str_pad = [function] (https://www.rdocumentation.org/packages/stringr/versions/1.5.0/topics/str_pad)
PorC_S = name of a source column (contains value '6' which we want add lead.zeros to)
6 = we want a final value to have exactly 6 places  
pad = "0" = empty places fill with "0"  

### Leading zeros -remove

```{r}

mutate(Mcislo = sub("^0+", "", Mcislo))
```

### ID and sequence along

```{r PORADI}

v <- c("raz", "dva", "tři", "čtyři", "pět") 
v <- as.data.frame(v) %>% 
  mutate(por = seq_along(v),
         porplus = seq_along(v)+1)
```

### Sequence in interval

```{r}
mutate(MinvcINT = mapply(function(MinvcOD, MinvcDO) paste(seq(MinvcOD, MinvcDO, by = 1), collapse = ";"), MinvcOD, MinvcDO)) %>% 
```


### Round

round(x, #) = zaokrouhleni na # pocet decimalnich cisel
ceiling(x) = rounds values up to nearest integer
floor(x) = rounds values down to nearest integer
trunc(x) = truncates (cuts off) decimal places from values

### Date

28138 -> 13.01.1977
pokud mame oba formaty, resime pomoci vnorenzch ifelsu takto:

```{r}
df <- source %>% 
  mutate(Mdat_ser = as.numeric(ifelse(str_detect(datumNabyti, "^.....$"), datumNabyti, NA)),
         Mdat_trans = format(as.Date(Mdat_ser, origin = "1899-12-30"), "%d.%m.%Y"),
         datnab22 = ifelse(str_detect(datumNabyti, "^.....$"), Mdat_trans, 
                           ifelse(str_detect(datumNabyti, "\\."), sub("\\.$", "", datumNabyti), datumNabyti)))
```

Mdat_ser = seriove datum
Mdat_trans = transformovane datum


\newpage

## S T R I N G S 

### Length

```{r STRING LENGTH}

mutate(Mpskp_len = nchar(pskup)) 
```

### Substring

```{r}

```


### Similarity

install.packages("RecordLinkage")
require("RecordLinkage")

levenshteinSim = funkce
autor_MUS, autor_AOPK = sloupce, ktere chci porovnat

```{r STRING SIMILARITY}

df %>% mutate(sim = levenshteinSim(autor_MUS, autor_AOPK))
```

### Detect case INsensitive

```{r STRING CASE SENS}

str_detect(MFN, regex("^FM\\s\\d+\\.jpg$", ignore_case = T))
```

### Separate

```{r SPLIT STRING}

# # Split name column into firstname and last name
# df[c('First Name', 'Last Name')] <- str_split_fixed(df$Name, ' ', 2)
#  
# # Rearrange columns and remove original name column
# df <- df[c('First Name', 'Last Name', 'State')]

separate(trida,into=c("trida_CZ", "trida_LAT") , sep=" ")

separate(MDRUH, into=c("MROD", "rest") , sep=" ", extra = "merge") # at first whitespace

separate(MDRUH, into=c("MROD", "rest") , sep=" ", extra = "merge", remove = F) # ponecha puvodni sloupec
```

### Keep text before

```{r}
  mutate(blMCELED = sub("\\/.*", "", fiMTREE)) # ponechej text do prvni /
```

### Extract word

```{r}
mutate(boMROD = word(druh, 1)) # vezmi 1. slovo ze sloupce druh
```

### Remove parentheses

```{r}

mutate(celed_LAT = gsub("[()]", "", celed_LAT))
```

### Remove whitespace

```{r RMV WHITE SPACE}

# package
require(stringr) # ::str_pad

# command
mutate(MMATERIAL = str_trim(MMAT_AGG)) 
```

str_trim = [function](https://stringr.tidyverse.org/reference/str_trim.html)

### Unite

Creates a string of multiple column values with NAs.

```{r}

unite(sys_tree_BIOL, c("kmen", "trida", "rad", "celed", "rod", "druh_Nomen"),
      sep = "/", na.rm = T, remove = FALSE) 
```

unite = [function](https://www.rdocumentation.org/packages/tidyr/versions/0.8.2/topics/unite)  
sys_tree_BIOL = name of a new column  
c(***) = list of columns  
sep = sign for separation  
na.rm = skip columns containing NA  
remove = keep or delete source columns  

### Seq_along with letters

```{r}
# lowercase
mutate(Msub = case_when(n()!=1 ~ letters[seq_along(porc)], TRUE ~ NA))

# uppercase
mutate(Msub = case_when(n()!=1 ~ LETTERS[seq_along(porc)], TRUE ~ NA))

```


\newpage

## D A T A F R A M E S

### e d a 

#### Any NA ?



### e n v i r o n m e n t

#### remove df with name pattern

remove all dataframes that contains "muz_" in their name

```{r}
rm(list = ls()[grepl("muz_", ls())])
```


### g e n e r a l

#### Replace "" with NA

s pipe jsem zatim cestu nenasla

```{r REPLACE "" WITH NA}

taxon[taxon == ""] <- NA
```

#### Replace NA with ""

```{r REPLACE NA WITH ""}

ou <- taxon %>%
  replace(is.na(.), "")

ou <- taxon %>%
  mutate_all(~ ifelse(is.na(.), "", .))
```

#### str_trim pro cely df

```{r STR_TRIM DF}

df %>% 
  mutate(across(where(is.character), str_trim))
```


#### mult column summary

```{r}
unique(df[,c('session','first','last')])

plyr::count(SAVE3[,c('zmena','duvod')]) # i s poctem
```


### c o l u m n s

#### Remove columns acc to pattern

```{r}

bot <- bot %>% 
  select(-contains("_sort"), -nvg_id)
```

#### Remove last n columns

```{r}

select(-mus_invc, -(last_col(offset = 34):last_col()))
```

#### Remove multiple colums

typicky nazvy sloupcu z biolibu

```{r}
nam <- colnames(bl_sp)
noquote(nam)
nam2 <- paste0("-", nam)
noquote(nam2)
nam3 <- paste0(nam2, collapse = ",")
nam3
```


#### Add prefix to column names

```{r}
colnames(kmen) <- paste0("kmen_", colnames(kmen))
```

#### Merge two columns with not complete information

column 1 has some NAs  
column 2 has also some NAs but in different rows than column 1
TASK: create one column with merged info

```{r}
#package
require(dplyr) # part of tidyverse

# command
mutate(kmen = coalesce(kmen_Nomen, trida_NadNomen))
```

coalesce = [function](https://www.rdocumentation.org/packages/dplyr/versions/1.0.10/topics/coalesce)

#### Relocate columns

```{r}
relocate(MINVC, .before = JineC_S) 
relocate(MINVCDO, .after = MINVC)

# more than 1, in this order

relocate(Mrada, Mcislo,  Mcislo0, Msub_a, .after = prirc) 

```

#### Shift columns 

```{r}
rows_to_change <- a$zmena == "N" # filter rows to be shifted
a[rows_to_change,3:12] <- a[rows_to_change,2:12] # 3:12 nova lokace, 2:12 stara lokace
a[rows_to_change,2] <- NA # puvodni lukaci (diru) nahradit NA
```


### r o w s

#### DF top 10 rows

```{r}

mus_tab_cvic <- mus_tab %>% 
  slice(1:10)
```

#### Random sample of n

```{r}

ch <- sample_n(df, 10) # n=10
```

#### Drop NA rows in spec col

```{r}

sez_int <- sez %>% 
  drop_na(MINVChorni)  
```

#### Filter with contains

```{r}

spo_ces <- ic %>% 
  filter(grepl("Další|Jiná|Lapidárium|Etnografická", zdroj))
```

#### Copy rows / multiply

1:nrow(df) = vsechny radky
, 1), ]) = jednou (tzn. ve vysledku dostaneme kazdy radek dvakrat)

```{r}

df <- rbind(df, df[rep(1:nrow(df), 1), ])
```

#### Duplicates

```{r}

#  https://community.rstudio.com/t/filter-out-all-rows-with-duplicate-values/41043/4

# Return names which have only a single row of data
data %>% 
  group_by(name) %>% 
  filter(n()==1)

# Return names which have more than one row of data
data %>% 
  group_by(name) %>% 
  filter(n() != 1)
```

\newpage

## R E G E X S

### extract all numbers

### extract all letters

### case sensitiviy

(?i)....

```{r}
str_detect(zpNabyti, "(?i)dar")
```

### replace more than one pattern

```{r}
str_replace_all(prirc, c("\\d+" = "", "\\/" = ""))
```



\newpage

## D A T A B A S E S

### Connect db and read table

```{r}
# package 
require(RODBC)

# general

dbname <- paste0("---cesta+jmeno-databaze.accdb") 
con <- RODBC::odbcConnectAccess2007(dbname)      
RODBC::sqlTables(con)
tab <- RODBC::sqlFetch(con, "---jmeno-tabulky---")

# old func example
dbname <- paste0(path_mdb, "botanika_test.mdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname)      # set the connection

RODBC::sqlTables(con)                          # list tables in database

kmen <- RODBC::sqlFetch(con, "data") # read specified table

path_mdb <- "M:/03 klienti/kraj moravskoslezsky/muzeum beskyd - MUZBE/02 muzBE konverze/zdrojdat/BIO/MB_PVO_Botanika/"

```

*.mdb = old MS ACCESS  
*.accdb = new MS ACCESS 


### List files from one or more folders

```{r}

# one
vypis <- as.data.frame(list.files("D:/MUZBE/Negativy/Image", full.names = T)) 

# more
vypis <- as.data.frame(list.files(c("D:/MUZBE/Obecné sbírky-4složky obrázků/Image_01",
                                    "D:/MUZBE/Obecné sbírky-4složky obrázků/Image_02",
                                    "D:/MUZBE/Obecné sbírky-4složky obrázků/Image_03",
                                    "D:/MUZBE/Obecné sbírky-4složky obrázků/Image_04"), 
                                  pattern = "jpg$|JPG$", full.names = T)) 

colnames(vypis)[1] <- "path"
```


## L O O P S

### @@@ Read all files from folder

1) make a list of files

```{r}
Lxls <- list.files(path_data, pattern = "*.xlsx")  # *.txt *.jpg ...
```

2) create expty dataframe and specify col namens

```{r}
df_filled <- data.frame()
# df_filled <- data.frame(matrix(ncol = 15, nrow = 0))

colnam <- c("prirc","rok_prir","dat_nabyt")
```
 
3) read files and bind them 

```{r}

for(i in Lxls) {
  a <- read_excel(paste0(path_data, i),
                  col_types = "text",
                  range = cell_cols("A:P"), na = " ") %>% # skip = 2 prepsano rangem
    slice(-1)
  names(a) <- colnam
  a$zdroj <- paste(i) # add source txt file name
  # print(count(a))
  df_filled <- rbind(df_filled, a)
  print(i)
  # rm(a)
}
```



## J O I N S

*inner_join* only keeps observations from x that have a matching key in y
*left_join()* keeps all observations in x.
*right_join()* keeps all observations in y.
*full_join()* keeps all observations in x and y.

[description and parameters](https://dplyr.tidyverse.org/reference/mutate-joins.html)

### Join only selected columns

```{r JOIN SEL COL}

left_join(biolib %>% dplyr::select('4nazevKategorie', '5nomen', '8nadrizenyTaxon'),
            by = c("MNOMEN" = "5nomen"), keep = T)
```


\newpage

# *REAL TASKS*

## Auto mus_tab -> csv

Je treba mit vytvoreny list policek, ktere maji byt ve vyslednem csv. Priklad: pro PK 'Lpk' obsahuje 105 nazvu poli pro CSV se 105 sloupci.

Zaroven je treba, aby se v pracovni tabulce nachazely prislusne sloupce se stejnym nazvem.

Funkce potom vyzobe a posklada vysledne CSV.

```{r}

# list sloupcu pro nove CSV

Lpk <- c("sbirka1", "podsb2", "cisrada3", "prirc4", "invc5", "lok6","prirpla7", "prirpla8", "prirpla9", "ozn10",
         "popis11", "rok12", "pocetks13", "obal14", "evid15", "mater16", "datace17", "stav18", "popstav19", "pozn20",
         "liter21", "datnab22", "zpnab23", "typ24","doklnab25", "mena26", "cennab27", "puvmaj28", "nabpozn29", "typ30",
         "armisto31", "arkontext32", "arakce33", "arvyzk34", "arlok35", "arku36", "arparc37", "arpoz38", "nalpozn39", "naladr40",
         "nalokr41", "nalobec42", "nalcast43", "nalmc44", "nalcis45", "nalul46", "nalorc47", "nalpsc48", "stat49", "artext50",
         "polpop51", "polCES52", "polks53", "poldopl1_54", " poldopl1_55", "poldopl2_56", "poldopl2_57", "poldopl3_58", "poldopl3_59", "archiv60",
         "acvsign61", "acvmark62", "acvpozn63", "vyrdat64", "vyrduv65", "vyrtyp66", "vyrdok67", "vyrmena68", "vyrcena69", "zpub70",
         "vyrCES71", "novaj72", "poznvyr73", "predchmaj74", "userins75", "datins76", "arkontext2_77", "arkontext3_78", "arkomp79", "arSJTSK80",
         "arcizivyz81", "arnazevvyz82", "arrokvyz83", "arfirmavyz84", "arvedoucivyz85", "rokakvi86", "prisp87", 
         "text1_88", "text2_89", "text3_90", "text4_91", "text5_92", 
         "priz1_93", "priz2_94", "priz3_95", "priz4_96", "priz5_97", 
         "datzap98", "text6_99", "text7_100", "text8_101", "text9_102", "text10_103", "porc104", "subc105")

# vytvorit prazdny dataframe pro zapis

n <- 32004 # pocet radku dle poctu zaznamu
imp <- "dfOST_akt_invc" # jmeno importu
   
mus_tab <- data.frame(smazme = 1:n) # prazdny dataframe

# tvorba CSV

for(i in Lpk){
  if(i %in% colnames(ee)){
    ano <- ee %>% select(all_of(i))
    mus_tab <- cbind(mus_tab, ano)
    print("jo")
  } else{
    ne <- data.frame(rep(NA, n))
    colnames(ne) <- i
    mus_tab <- cbind(mus_tab, ne)
    print("houby")}
}

mus_tab <- mus_tab %>% select(-smazme) 

```



## Save mod_tab a mus_tab

imp = importni rozhrani 
f2s = file to save

```{r SAVE CSV AND OPER FILE}

# finalni csv

f2s <- paste0(path_csv, projekt, "_import_", imp, ".csv")

if(file.exists(f2s)){
  print("You've been here before!")
  rm(f2s)
} else {
  write.table(mus_tab, file = f2s,
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="cp1250")
  print("Import csv  W R I T T E N  !")
  rm(f2s)
  }

# upravena tabulka

mod2s <- paste0(path_csv, projekt, "_modif_", imp, ".csv")

if(file.exists(mod2s)){
  print("You've been here before!")
  rm(mod2s)
} else {
  write.table(mod_tab, file = mod2s, 
            quote = T, row.names = F, 
            sep = ";", dec = ",",              # desetinny oddelovac musi byt carka
            na = "", fileEncoding="cp1250")
  print("Operational file  S A V E D  !")
  rm(mod2s)
  }
```


## Uprava dataci

```{r DERIVE VZNIKPTU FROM SBIRKY, warning = F, eval = F}

dopr_tab <- sbirky %>% 
  select(MINVC, Autor_S, Role_S, Datace_S, MistoVz_S) %>% 
  mutate(MEMPTY = "") %>% 
  mutate(MDATACE = Datace_S) %>% 
  # yyyy
  mutate(MDAT_Y = ifelse(str_detect(MDATACE, "_00_00"), 
                         str_replace_all(MDATACE, "_00|-00", ""), NA)) %>% 
  # mm.yyyy
  # mutate(MDAT_YM = ifelse(str_detect(MDATACE,
  #                                    "^([0-9]{4})[_]([0-9]{1})([1-9]{1})[_]00$"),
  #                                   # format(anydate(substring(MDATACE, 1, 7), "%m.%Y")), ""))
  #                                   # substring(format(anydate(MDATACE), "%m.%Y"), 1, 7), ""))
  #                                   format(anydate(MDATACE), "%m.%Y"), ""))
  # mutate(MDAT_YM = format(anydate(MDATACE), "%m.%Y")) %>%
  mutate(MDAT_YM = substring(ifelse(str_detect(MDATACE,
                                             "^([0-9]{4})[_]([0-9]{1})([1-9]{1})[_]00$"),
                                  MDATACE, ""), 1, 7)) %>%
  mutate(MDAT_YM = format(anydate(MDAT_YM), "%m.%Y")) %>%
  # dd.mm.yyyy
  mutate(MDAT_YMD = ifelse(str_detect(MDATACE,
                          "^([0-9]{4})[_]([0-9]{2})[_]([0-9]{2})$"), 
                           format(anydate(MDATACE), "%d.%m.%Y"), NA)) %>% 
  # yy leta ss stoleti
  mutate(MDAT_LETA = ifelse(str_detect(MDATACE, "^([0-9]{4})[_]XX[_]([0-9]{2})$"),
                                    paste0(substring(dopr_tab$MDATACE, 9, 9), "0. léta ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>% 
  # poloviny
  mutate(MDAT_2 = ifelse(str_detect(MDATACE, "^00\\d\\d[_]\\d2[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". polovina ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>%
  # tretiny - NEVYSKYTUJE SE
  mutate(MDAT_3 = ifelse(str_detect(MDATACE, "^00\\d\\d[_]\\d3[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". třetina ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>%
  # poloviny - NEVYSKYTUJE SE
  mutate(MDAT_4 = ifelse(str_detect(MDATACE, "^00\\d\\d[_]\\d3[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". čtvrtina ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>% 
  # tretina roku
  mutate(MDAT_3r = ifelse(str_detect(MDATACE, "^\\d([1-9]{2})\\d[_]\\d3[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". třetina roku ",
                                           substring(dopr_tab$MDATACE, 1, 4)), NA)) %>% 
  # DATACE FINAL
  mutate(MDATACE_FIN = coalesce(MDAT_Y, MDAT_YM, MDAT_YMD, MDAT_LETA, MDAT_2, MDAT_3r)) %>% 
  mutate(MDATACE_FIN = coalesce(MDATACE_FIN, MDATACE))
```

## Tvorba inventarniho cisla

```{r MINVC}

sbirky <- read_excel(paste0(path_data, "/Sbirky.xlsx")) %>% 
  
  select(-MSUBOD, -MSUBDO, -MINVC, -MINVCDO) %>%    # get rid of .accdb columns
  
  mutate(MPORC = str_pad(PorC_S, 6, pad = "0"),
         MPORCDO = str_pad(HPorC_S, 6, pad = "0"),
         MSUB = ifelse(Lomeni_S == "_", NA, Lomeni_S),
         MSUBDO = HLomeni_S, 
         MINVC = case_when(is.na(MSUB) ~ paste0(Rada_S, "-", MPORC),
                             TRUE ~ paste0(Rada_S, "-", MPORC, "/", MSUB)),
         MINVCDO = case_when(is.na(MSUBDO) ~ "",
                             TRUE ~ paste0(Rada_S, "-", MPORCDO, "/", MSUBDO))) %>% 
  # select(-MPORC, -MPORCDO) %>%
  relocate(MINVC, .before = JineC_S) %>% 
  relocate(MINVCDO, .after = MINVC) 

# one step

MINVC = paste0("M", str_pad(invc, 5, pad = "0"))

```

## Taxon -> rod, druh, autor

```{r SPLIT TAXON}

separate(taxon, into=c("Mrod", "Mdruh", "Mautor") , sep=" ", extra = "merge", remove = F)
```

## Check import KD

```{r KD IMPORT CHECK}

mus_imp <- read_excel(paste0(path_kd, v, "/KontextovyDokument_export (1).xlsx")) %>%  
  mutate(MFN = sub("_original", "", URL))

jpg_cs <- jpg %>% mutate(MFN = ifelse(str_detect(MFN, "JPG"), sub("JPG", "jpg", MFN), MFN))

imp_check <- left_join(jpg_cs, mus_imp, by = c("MFN" = "MFN"), keep = T)

na <- imp_check %>% filter(is.na(Číslo))

```


\newpage

# *NEED TO COVER*

dopr_tab <- sbirky %>% 
  select(MINVC, Technika_S) %>% 
  mutate(MEMPTY = "") %>% 
  mutate('2sbirkaCisloEvidInt' = sbirka) %>% 
  mutate('3podsbirkaCislo' = podsbirka) %>% 
  mutate(MTECH = Technika_S) %>% 
  separate(MTECH,into=c("MTECH1", "MTECH2", "MTECH3", "MTECH4", "MTECH5"),
           sep=",(?! tuší)") %>% # separator is ',' BUT ONLY IF NOT FOLLOWED BY ' tuší' -> solves a single case of 'kresba tužkou, tuší a bělobou' which should remain as one entry
  pivot_longer(
    cols = MTECH1:MTECH5, 
    names_to = "MTECH_POR", # poradi techniky
    values_to = "MTECH_AGG",
    values_drop_na = TRUE) %>% 
  mutate(MTECHNIKA = str_trim(MTECH_AGG)) %>% # remove white space
  select(-MTECH_AGG) %>% 
  mutate(MTECH_POR = substr(MTECH_POR, 6, 6))
  
//
  
  
plant <- db_tab %>% 
  
   separate(celed,into=c("celed_LAT", "celed_CZ") , sep=" ") %>% 
   separate(druh,into=c("rod_LAT", "druh_LAT", "rest"), sep = "\\s(?!x )", 
            extra = "merge", remove = FALSE) %>% 
   mutate(druh_LAT = paste(rod_LAT, druh_LAT),
          celed_CZ = gsub("[()]", "", celed_CZ),
          autor = ifelse(str_detect(rest, "^[[:upper:]]") | str_detect(rest, "^\\W"), rest,""),
          druh_LAT = gsub("x$", "", druh_LAT))

//

 mutate_all( ~ str_replace_all(., pattern = ";", replacement = " ")) %>% # removes semicolons
  mutate_all( ~ str_replace_all(., pattern = "[\r\n]", replacement = " ")) #removes paragraphs
  